Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-03 paper/paper.tex
pass on intro

This paper focuses on rule-based \nids{}, which are very popular in
industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
\emph{rule-based intrusion detector}\footnote{\aka\ signature-based
  intrusion detector.} checks if the network traffic matches a fixed
set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
of Suricata~\cite{suricata}, a popular open-source \nids{}. This
particular rule prescribes a method to capture an Active
Reconnaissance attack for discovering which ports of a publicly
accessible machine are open for connection (see
Section~\ref{sec:active-recon}). Rulesets exist for Snort~\cite{snort}
and Suricata~\cite{suricata}, providing protection against existing
attacks. Unfortunately, attackers actively look for new mechanisms to
exploit systems. For example, Figure~\Fix{plot} shows the increase in
number of rules added to the ``Emerging Threats Open Ruleset''---a
popular public data set of \suri\ rules---in the past 12 months.

@@

This paper proposes \tname{}\footnote{Abbreviation for
  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
technique that uses machine intelligence to automatically synthesize
rules for rule-based \nids. Our goal is to facilitate the creation of
rules and, consequently, help system administrators to harden the
protection of \nids\ to network attacks. \tname{} synthesizes rules
from malicious and benign traffic and from examples of correct rules.

>>>>

This paper focuses on rule-based \nids{}, which are very popular in
industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
\emph{rule-based intrusion detector}\footnote{\aka\ signature-based
  intrusion detector.} checks if the network traffic matches a fixed
set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
of Suricata~\cite{suricata}, a popular open-source \nids{}. This
particular rule prescribes a method to capture an Active
Reconnaissance attack for discovering IPS of hosts in a network (see
Section~\ref{sec:active-recon}). Rulesets exist for Snort~\cite{snort}
and Suricata~\cite{suricata}, providing protection against existing
attacks. Unfortunately, attackers actively look for new mechanisms to
exploit systems. For example,
Figure~\ref{fig:distribution-rules-per-month} shows the number of
rules added per month in the year of 2019 to the ``Emerging Threats
Open Ruleset'', a popular public data set of \suri\ rules. We used the
``creation-date'' field associated with each entry in the ruleset. The
data provides some evidence that new rules are created often.

@@

This paper proposes \tname{}\footnote{Abbreviation for
  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
technique that uses machine intelligence to automatically synthesize
rules for rule-based \nids. Our goal is to facilitate the creation of
rules and, consequently, help system administrators to harden the
protection of \nids\ to network attacks. \tname{} synthesizes rules to
protect against a given attack from 1) malicious traffic (\ie{},
examples of the attack to protect against), 2) benign traffic, and 3)
trusted rules.
